Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-14 paper/sec-mva.tex
All comments implemented

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.90\textwidth]{plots/bst_nnarch_noPU.pdf}
  \vspace{-2cm}
  \caption{\small Architecture of the Artificial
    Neural Network (ANN)
    used for the analysis of the
    boosted
    category, with $N_{\rm var}=21$ input variables and thus
    the same number of neurons
  in the first layer.
  %
  The color code in the neuron connections (the weights) is a heat map obtained
  at the end of the GA training,
  with red indicating larger values and black smaller values.
}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

We now have all the
information needed to determine a suitable
value of the MVA output cut $y_{\rm cut}$ in each
of the three categories.
%
These values can be determined from the maximisation of $S/\sqrt{B}$,
ensuring that the number of signal events $N_{\rm ev}$
expected at the HL-LHC is not too small.
%
In  addition, we require
that the number of MC events used to define the signal
category (events with $y_i \ge y_{\rm cut}$)
is large enough to avoid the biases and statistical
fluctuations associated to a small training sample.
%
In Table~\ref{table:cutflowMVA} where we indicate
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC as well as $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
    \hline
    \multicolumn{6}{|c|}{No PU} \\
    \hline
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 1070 & $7.6\cdot 10^5$  & 1.2  & $1.4\cdot 10^{-3}$  \\
    &  $y_{\rm cut}=0.82$ & 790  & $2.2\cdot 10^4$   & 5.4  & 0.034 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  & 670   & $5.3\cdot 10^5$
    & 0.9 & $1.2\cdot 10^{-3}$ \\
    &  $y_{\rm cut}=0.80$ & 360  & $5.5\cdot 10^3$  & 4.8 & 0.06\\
    \hline
    \hline
      \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  & 3700 &  $1.5\cdot 10^{7}$ &  1.0 &$3\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.65$ & 1300  & $8.3\cdot 10^{4}$ & 4.5 & 0.01 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA results, for the optimal value of the
    ANN discriminant $y_{\rm cut}$ in the three categories, compared with the
    corresponding
    pre-MVA results $y_{\rm cut}=0$.
    %
    We indicate the number of signal and
    background events
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    the signal significance $S/\sqrt{B}$ and
    the signal over background ratio $S/B$.
    %
    The pre-MVA results ($y_{\rm cut}=0$) corresponds to row C2 in
    Table~\ref{tab:cutflow_noPU_1}.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/roc_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/nev2_SKPU80.pdf}
\caption{\small Same as Fig.~\ref{fig:nev2} in the
case of events with PU, for
 $\la n_{\rm PU}\ra=80$ 
  and SK subtraction.
}
\label{fig:nev2_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_SKPU80.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_SKPU80.pdf}
\caption{\small 
Same as Fig.~\ref{fig:sb_mva} in the
case of events with PU, for
 $\la n_{\rm PU}\ra=80$ 
  and SK subtraction.
}
\label{fig:sb_mva_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_SKPU80.pdf}
\vspace{-0.5cm}
\caption{\small
Same as Fig.~\ref{fig:nnweights} in the
case of events with PU, for
 $\la n_{\rm PU}\ra=80$ 
  and SK subtraction.
}
\label{fig:nnweights_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

In Table~\ref{table:cutflowMVA_fakes} we compare,
after the MVA is applied,
the total
number of background events, $N_{\rm ev}^{\rm tot}$,
    expected at the HL-LHC, with
     $N_{\rm ev}^{\rm 4b}$, the corresponding number
    from the QCD $4b$ component, in the cases with and without
    PU.
    %
    The MVA has been trained to the total background sample,
    though differences
    in the kinematic distributions of the $4b$ and $2b2j$ processes are small,
    see Fig.~\ref{fig:histoBack}.
    %
    The major improvement of eliminating the contamination
    of light and charm jet mis-identification is in the boosted category.

    In the case of PU+SK, combining the three categories but assuming that
    the only relevant background is the irreducible QCD $4b$ component,
    we obtain
    \be
\lp \frac{S}{\sqrt{B_{\rm 4b}}}\rp_{\rm tot} \simeq 7.4(2.4) \, ,\quad
\mathcal{L}=3000 (300)\,{\rm fb}^{-1}\, ,
\ee
Therefore, a discovery of Higgs pair production at the
HL-LHC might be possible if the
mis-identification of light and charm jets as
$b$-jets can be  reduced.
%
With these assumptions, and provided
the signal efficiency can be further enhanced,
this measurement might even be feasible
by the end of Run II.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{|c|c|c|c||c|c||c|c|}
        \hline
        Category  &    &  $N_{\rm ev}^{\rm tot}$  &  $N_{\rm ev}^{\rm 4b}$   &
        $S/\sqrt{B_{\rm tot}}$ & $S/\sqrt{B_{\rm 4b}}$  
        &  $S/B_{\rm tot}$ & $S/B_{\rm 4b}$\\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  no PU  & $2.2\cdot 10^4$  & 4800    & 
      5.4 &  11.6 & 0.034 & 0.16 \\
    & PU+SK & 7300  &  1600  &  2.9  & 6.2 &  0.034 & 0.16 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  no PU   & $5.5\cdot 10^3$  & 2000    &
    4.8  &  7.9 &  0.06  &  0.16 \\
    & PU+SK  & 2600   &  1000  & 1.1  & 1.8 & 0.02 & 0.05 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &   no PU  & $8\cdot 10^4$   &
    $3.7\cdot 10^4$
    & 4.5  & 6.7  & 0.01 & 0.02 \\
    & PU+SK  &  $3\cdot 10^4$   &   $1.3\cdot 10^4$ & 2.5    & 3.7  & 0.015 & 0.03 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA number of total background events, $N_{\rm ev}^{\rm tot}$,
    expected at the HL-LHC, compared
    with $N_{\rm ev}^{\rm 4b}$,
    the number of QCD $4b$ background events.
    %
    We also indicate the corresponding signal significance and signal over background
    ratios.
        \label{table:cutflowMVA_fakes}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.90\textwidth]{plots/bst_nnarch_noPU.pdf}
  \vspace{-2cm}
  \caption{\small Architecture of the Artificial
    Neural Network (ANN)
    used for the analysis of the
    boosted
    category, with $N_{\rm var}=21$ input variables and thus
    the same number of neurons
  in the first layer.
  %
  The color code in the neuron connections (the weights) is a heat map obtained
  at the end of the GA training,
  with red indicating larger values and black indicating smaller values.
}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

We now have all the
information needed to determine a suitable
value of the MVA output cut $y_{\rm cut}$ in each
of the three categories.
%
These values can be determined from the maximisation of $S/\sqrt{B}$,
ensuring that the number of signal events $N_{\rm ev}$
expected at the HL-LHC is not too small.
%
In  addition, we require
that the number of MC events used to define the signal
category (events with $y_i \ge y_{\rm cut}$)
is large enough to avoid the biases and statistical
fluctuations associated to a small training sample.
%
In Table~\ref{table:cutflowMVA} where we quote
the value of the optimal ANN output
cut in each category,
the number of signal and background events $N_{\rm ev}$ expected
at the HL-LHC as well as $S/\sqrt{B}$ and $S/B$.
%
For reference, we also include the results at the end of
the cut-based
analysis.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|l|c|c|c|c|}
    \hline
    \multicolumn{6}{|c|}{No PU} \\
    \hline
    \hline
    Category  &   &  $N_{\rm ev}$ signal &  $N_{\rm ev}$ back  &  $S/\sqrt{B}$ & $S/B$ \\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  $y_{\rm cut}=0$  & 1070 & $7.6\cdot 10^5$  & 1.2  & $1.4\cdot 10^{-3}$  \\
    &  $y_{\rm cut}=0.82$ & 790  & $2.2\cdot 10^4$   & 5.4  & 0.034 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  $y_{\rm cut}=0$  & 670   & $5.3\cdot 10^5$
    & 0.9 & $1.2\cdot 10^{-3}$ \\
    &  $y_{\rm cut}=0.80$ & 360  & $5.5\cdot 10^3$  & 4.8 & 0.06\\
    \hline
    \hline
      \multirow{2}{*}{Resolved} &  $y_{\rm cut}=0$  & 3700 &  $1.5\cdot 10^{7}$ &  1.0 &$3\cdot 10^{-4}$ \\
    &  $y_{\rm cut}=0.65$ & 1300  & $8.3\cdot 10^{4}$ & 4.5 & 0.01 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA results, for the optimal value of the
    ANN discriminant $y_{\rm cut}$ in the three categories, compared with the
    corresponding
    pre-MVA results $y_{\rm cut}=0$.
    %
    we quote the number of signal and
    background events
    at the HL-LHC with $\mathcal{L}=3$ ab$^{-1}$,
    the signal significance $S/\sqrt{B}$ and
    the signal over background ratio $S/B$.
    %
    The pre-MVA results ($y_{\rm cut}=0$) corresponds to row C2 in
    Table~\ref{tab:cutflow_noPU_1}.
    \label{table:cutflowMVA}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
    \includegraphics[width=0.49\textwidth]{plots/roc_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/nev2_SKPU80.pdf}
\caption{\small Same as Fig.~\ref{fig:nev2} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$ using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:nev2_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_SKPU80.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_SKPU80.pdf}
\caption{\small 
Same as Fig.~\ref{fig:sb_mva} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$
using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:sb_mva_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_SKPU80.pdf}
\vspace{-0.5cm}
\caption{\small
Same as Fig.~\ref{fig:nnweights} in the
case of events with PU, for
$\la n_{\rm PU}\ra=80$
using
{\tt SoftKiller} for PU subtraction.
}
\label{fig:nnweights_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

In Table~\ref{table:cutflowMVA_fakes} we compare,
after the MVA is applied,
the total
number of background events, $N_{\rm ev}^{\rm tot}$,
    expected at the HL-LHC, with
     $N_{\rm ev}^{\rm 4b}$, the corresponding number
    from the QCD $4b$ component, in the cases with and without
    PU.
    %
    The MVA has been trained to the total background sample,
    though differences
    in the kinematic distributions of the $4b$ and $2b2j$ processes are small,
    see Fig.~\ref{fig:histoBack}.
    %
    The major improvement of eliminating the contamination
    of light and charm jet mis-identification is for the boosted and resolved categories.
%
    In the case of PU+SK, combining the three categories but assuming that
    the only relevant background is the irreducible QCD $4b$ component,
    we obtain
    \be
\lp \frac{S}{\sqrt{B_{\rm 4b}}}\rp_{\rm tot} \simeq 5.3~(1.8) \, ,\quad
\mathcal{L}=3000~(300)\,{\rm fb}^{-1}\, ,
\ee
Therefore, a discovery of Higgs pair production at the
HL-LHC might be possible if the
mis-identification of light and charm jets as
$b$-jets can be  reduced.
%
With these assumptions, and provided
the signal efficiency can be further enhanced,
this measurement might even be feasible
by the end of Run II.
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{|c|c|c|c||c|c||c|c|}
        \hline
        Category  &    &  $N_{\rm ev}^{\rm tot}$  &  $N_{\rm ev}^{\rm 4b}$   &
        $S/\sqrt{B_{\rm tot}}$ & $S/\sqrt{B_{\rm 4b}}$  
        &  $S/B_{\rm tot}$ & $S/B_{\rm 4b}$\\ 
    \hline
    \hline
    \multirow{2}{*}{Boosted} &  no PU  & $2.2\cdot 10^4$  & $1.5\cdot 10^4$     & 
      5.4 &  6.5 & 0.034 & 0.05 \\
    & PU+SK & 7300  &  500  &  2.9  & 3.6 &  0.034 & 0.04 \\
    \hline
    \hline
    \multirow{2}{*}{Intermediate} &  no PU   & 4500  & 2000    &
    4.8  &  5.3 &  0.06  &  0.08 \\
    & PU+SK  & 2600   &  1500  & 1.1  & 1.2 & 0.02 & 0.03 \\
    \hline
    \hline
    \multirow{2}{*}{Resolved} &   no PU  & $8\cdot 10^4$   &
    $4.5\cdot 10^4$
    & 4.5  & 6.6  & 0.01 & 0.03 \\
    & PU+SK  &  $3\cdot 10^4$   &   $1.3\cdot 10^4$ & 2.5    & 3.7  & 0.015 & 0.03 \\
    \hline
      \end{tabular}
  \caption{\small Post-MVA number of total background events, $N_{\rm ev}^{\rm tot}$,
    expected at the HL-LHC, compared
    with $N_{\rm ev}^{\rm 4b}$,
    the number of QCD $4b$ background events.
    %
    We also indicate the corresponding signal significance and signal over background
    ratios.
        \label{table:cutflowMVA_fakes}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
